---
title: "R Notebook"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(leaflet)
library(sf)
library(readxl)
library(DT)
library(plotly)
library(broom)
library(tidycensus)
```

```{r}
senate_counties <- read_xlsx("Statewide Results.xlsx", sheet = 1)
```
This loads the data.

```{r}
senate_counties <- read_xlsx("Statewide Results.xlsx", sheet = 1, range = "B7:E63")
```
This loads the data.

```{r}
glimpse(senate_counties)
```

```{r include=FALSE}
census_api_key("415c7b9b938b2032bb48c2a0203b0059ceece209", overwrite = TRUE, install = TRUE)

```

```{r include=FALSE}
mt_counties <- get_acs(geography = "county",
                       variables = "B01003_001",
                       state = "MT",
                       geometry = TRUE) 
```


This shows a preview of data.

```{r}
senate_counties <- senate_counties %>% 
  rename(Republican = "MATT ROSENDALE\r\nRepublican") %>% 
  rename(Democrat = "JON TESTER\r\nDemocrat") %>% 
  rename(Libertarian = "RICK BRECKENRIDGE\r\nLibertarian")
```

```{r}

senate_counties <- senate_counties %>% 
  mutate(total_votes = Republican + Democrat + Libertarian) %>% 
  mutate(Repub_advantage = Republican/total_votes - Democrat/total_votes) %>% 
  mutate(Repub_advantage = round(Repub_advantage*100, 1))
  senate_counties %>% 
  arrange(-Repub_advantage)
```
This is a table showing the the votes for the candidates by county a long with the republican advantage of each county.

```{r}
mt_counties <- get_acs(geography = "county",
                       variables = "B01003_001",
                       state = "MT",
                       geometry = TRUE) 
```
This gets data for state.
```{r}

senate_counties[25, "County"] <- "Lewis and Clark"              # Changes  "&" "and"

mt_counties <- mt_counties %>% 
  mutate(County = gsub(" County, Montana", "", NAME)) %>%      # Removes unnecessary words
  rename(Population = estimate)                                # Renames the 'estimate' to 'Population'
```
This is just replacing and with & from two mismatched titles.
```{r}
senate_election <- mt_counties %>% 
  full_join(senate_counties)
```
This joins the data from mt counties with the senate race data.
```{r}
senate_election %>%
  as_tibble() %>% 
  select(County, Population, Democrat, Republican, Libertarian, total_votes, Repub_advantage) %>% 
  datatable()
```
This is a table showing the counties with thier population and the votes per candidate with the republican advantage.
```{r}

vote_colors <- colorNumeric(palette = "viridis", domain = senate_election$Repub_advantage)

senate_election %>%
  leaflet() %>% 
  addTiles() %>%
  addPolygons(weight = 1,
              fillColor = ~vote_colors(Repub_advantage), 
              label = ~paste0(County, ", Republican advantage = ", Repub_advantage),
              highlight = highlightOptions(weight = 2)) %>% 
  setView(-110, 47, zoom = 6) %>% 
  addLegend(pal = vote_colors, values = ~Repub_advantage)

```
This shows the republican advantage by county. The lighter or more toward yellow indicates a stronger advantage.

```{r}
senate_election %>%
  plot_ly(x = ~Population, 
          y = ~Repub_advantage,
          hoverinfo = "text", 
          text = ~paste("County:", 
                        County, "<br>", 
                        "Population: ", Population, "<br>", 
                        "Republican advantage: ", Repub_advantage)) %>% 
  add_markers(marker = list(opacity = 0.7)) %>%
  layout(title = "Predicting Republican Vote Advantage from Population, by County",
         xaxis = list(title = "County population"),
         yaxis = list(title = "Republican vote advantage"))
```
This is a plotly graph idicating a negative correllation.
```{r}
pop_model <- lm(Repub_advantage ~ Population, data = senate_election)

```
This is a linear regression.
```{r}
summary(pop_model)
```
This is a summary of the regression results.
```{r}
tidy(pop_model)
glance(pop_model)
```
This shows the information in a cleaner way.

```{r}
senate_election %>%
  plot_ly(x = ~Population, 
          y = ~Repub_advantage,
          hoverinfo = "text", 
          text = ~paste("County:", 
                        County, "<br>", 
                        "Population: ", Population, "<br>", 
                        "Republican advantage: ", Repub_advantage)) %>% 
  add_markers(showlegend = F, marker = list(opacity = 0.7)) %>%
  layout(title = "Predicting Republican Vote Advantage from Population, by County",
         xaxis = list(title = "County population"),
         yaxis = list(title = "Republican vote advantage")) %>%
  add_lines(y = ~fitted(pop_model))
```
This predicts the republican vote advantage from population and by county.

```{r echo=TRUE}
senate_election <- senate_election %>% 
  mutate(Longitude = as_tibble(st_coordinates(st_centroid(senate_election$geometry)))$X) %>% 
  mutate(Latitude = as_tibble(st_coordinates(st_centroid(senate_election$geometry)))$Y)
```
```{r}
senate_election %>%
  leaflet() %>% 
  addTiles() %>%
  addPolygons(weight = 1) %>% 
  setView(-110, 47, zoom = 6) %>% 
addCircleMarkers(~Longitude, ~Latitude)
```
This is just another graph showing the center of counties.
```{r}
senate_election %>%
  plot_ly(x = ~Longitude, y = ~Repub_advantage) %>% 
  add_markers()
```
This is a plotly graph of the longitute and republican vote indicating a positive correlation that the further east the more votes for republican.
```{r}
longitude_lm <- lm(Repub_advantage ~ Longitude, data = senate_election)
tidy(longitude_lm)
glance(longitude_lm)
```
This is a regression predicting republican vote from longitude.
```{r}
senate_election %>% 
  plot_ly(x = ~Longitude, 
          y = ~Repub_advantage,
          hoverinfo = "text", 
          text = ~paste("County:", County, "<br>", "Longitude: ", Longitude, "<br>", "Republican advantage: ", Repub_advantage)) %>% 
  add_markers(marker = list(opacity = 0.7), showlegend = F) %>%
  layout(title = "Predicting Republican Vote Advantage from Longitude, by County",
         xaxis = list(title = "County longitude"),
         yaxis = list(title = "Republican vote advantage")) %>% 
  add_lines(y = ~fitted(longitude_lm))
```
This is a plotly graph predicting the republican vote advantage from longitude, by county.
```{r}
multiple_lm <- lm(Repub_advantage ~ Population + Longitude, data = senate_election)
tidy(multiple_lm)
glance(multiple_lm)
```
This is a multiple regression, predicting the Republican advantage from both Population and Longitude at the same time.
```{r}
senate_election %>% 
  plot_ly(x = ~Longitude, y = ~Population, z = ~Repub_advantage, 
          text = ~County, hoverinfo = "text") %>% 
  add_markers(opacity = .7, showlegend = F)
```
This is a 3D scatterplot showing the republican advantage, population, and votes per longitudinal data.































































